On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}

### 2020-04-07 engineering.tex
partial review of engineering section

Several standards mandates a secure-by-design approach in which 
cybersecurity shall be considered at the very early stages of the 
design process. For example,
\begin{enumerate}[noitemsep]
	\item DO-326A -- ``Airworthiness Security Process Specification''
		requires a cybersecurity risk assessment of the design and is
		the ``are the only Acceptable Means of Compliance (AMC) by FAA
		\& EASA for aviation cyber-security airworthiness certification,
		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
		Control System (ICS) Security''
	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
		for Cyber-Physical Vehicle Systems'' defines ``set of
		high-level guiding principles for Cybersecurity as it relates
		to cyber-physical vehicle systems'' and states that
		``incorporate Cybersecurity into cyber-physical vehicle systems
		from concept phase through production, operation, service, and
		decommissioning''
\end{enumerate}
Companies who develop CPS (e.g. aerospace systems or elevator systems), must adhere
to strict regulations and quality assurances processes. Introducing tools and procedures 
in their processes requires a detailed and justified overview on how those
tools and procedures can be used. In Figure~\ref{fig:spdl}, we provide an overview
of a process that consider cybersecurity in the specification, design, and implementation
stages. The overall process start with the specification of the architectural (both physical
and functional) requirements from which the Weaknesses are automatically identified.
Based on the Assets, architectures, and the Weaknesses the cybersecurity risk is
calculated based on the $\abftheory$-theory. An iterative process between 
the user (e.g. the engineer) and the \secramod module of the \abftool allows the user to move to the design phase
when the risk level is considered acceptable. The design phase starts by automatically mapping
the Weaknesses into Vulnerabilities which are, in turn, fed into the \designverifmod
module along with the final specification. The Module returns a list of Vulnerabilities and
potential Mitigations. The user then procees in implementing the CPS and the SW/HW choices
can be tested by the \atgmod.

@@

In order to explain in details the SPDL, we use the following running example.
\begin{example}
	As depicted in Figure~\ref{fig:usecase}, the use case of the running example
	(a small part of a subprocess of the SwAT testbed \autocite{Mathur2016swat}), shows
	\begin{itemize}[noitemsep]
		\item a tank that is filled with raw water coming from the inlet pipe 
		\item a motorized valve (a valve with an actuator) that regulates that
			\begin{itemize}
				\item if opened, allows the raw water to flow into the tank
				\item if closed, stops the raw water to flow into the tank
			\end{itemize}
		\item a sensor that reads the meters of raw water in the tank
		\item a PLC controller, connected to the sensor and the actuator, that receives the readings from the sensor and regulates the quantity of water in the tank by communicating a change in the state of the actuator based on the following logi
			\begin{itemize}
				\item if the readings from the sensor reports a water level in the tank above (or equal) to 10m, the actuator shall close the valve
				\item if th readings are below 10m the actuator shall open the valve.
			\end{itemize}
	\end{itemize}

@@

\subsection{Specification -- Elicitation of Security Requirements}\label{sec:properties}
\begin{figure}[t]
	\centering
	\includegraphics[width=\textwidth]{spec.pdf}
	\caption{Specification -- Running Example}
	\label{fig:spec}
\end{figure}
The first step of the SPDL is the definition of the requirements and then, for
the purpose of this paper of the (i) security requirements, (ii) architectural
(physical and functional) requirements.  Our focus in mainly on security
requirements but they are necessarily intertwined with the architectural
requirements, as we will show afterwards in this section.  A complete study on
how to properly define the architectural requirements is outside the scope of
this paper so we propose a method that is convenient for the purpose of our
assessment.

\Paragraph{System Requirements}
In our running example, there is a system (a CPS) composed by $5$ main
subsystem which we consider atomic (i.e. not themself composed by subsystems)
and then agents: the tank, the sensor, the controller, the actuator, and the
motorized valve (valve from now on).  The whole specification (which we now
describe) of the physical and functional architecture is depicted in
Figure~\ref{fig:spec}.  In the use case, the communication flow is as follows:
\begin{enumerate}
	\item the sensor reads the level of water in the tank ($\text{tank}\rightarrow \text{sensor}$) \footnote{this communication depends on the technology in use for the readings. We assume a unidirectional communication without loss of generality}
	\item the sensor communicates the readings to the controller ($\text{sensor}\rightarrow \text{controller}$)
	\item the controller calculates the valve state and communicates it to the actuator ($\text{controller}\rightarrow \text{actuator}$)
	\item the actuator translates (digital to analog) the states received and communicates\footnote{we do not distinguish, for the sake of simplicity between different types of channels and assume that communicating to the valve produces the expected change of physical state} it to the valve
\end{enumerate}
Each agent has an input or output port from which it communicates with other
agents.  It is also composed by the functional blocks that defines its
functional architecture and physical boundaries that defines its physical
architecture, as follows.
\begin{enumerate}[start=1, label={req\arabic*)}]
	\item the tank has no functional architecture, it is a purely physical component which has the only purpose of containing water
	\item the communication is unidirectional from the tank to the input port of the sensor
	\item the sensor has an input-port that receives only incoming communication from the tank
	\item the sensor has two functional blocks
		\begin{enumerate}[label={req4.\arabic*)}]
			\item read\_water\_level, accepts inputs only from the input-port and outputs a new reading every $t$ seconds
			\item comm\_water\_level, accepts inputs only from the read\_water\_level and communicates the readings to the output-port
		\end{enumerate}
	\item the sensor has an output-port that only outputs the messages generated by the comm\_water\_level to the input-port of the controller
	\item the controller has in input-port that only receives incoming communications from the sensor and sends the messages to the functional architecture
	\item the functional architecture of the controller is composed by three functional blocks
		\begin{enumerate}[label={req7.\arabic*)}]
			\item get\_water\_level, accepts inputs only from the input-port and outputs the water level to the main control function
			\item control, calculate the next state of the valve based on the inputs receive from the get\_water\_level
			\item send\_state, receives the output of the control function and communicates it to the output-port
		\end{enumerate}
	\item the controller has an output-port that only outputs the messages generated by the send\_state to the input-port of the actuator 
	\item the actuator has an input-port that receives only incoming communication from the controller 
	\item the functional architecture of the actuator is composed by two functional blocks
		\begin{enumerate}[label={req10.\arabic*)}]
			\item read\_change\_state, accepts inputs only from the input-port and outputs the next state of the valve to the close\_open\_valve function
			\item close\_open\_valve, converts the digital next state received from the read\_change\_state an analog signal and communicates it the output-port
		\end{enumerate}
	\item the valve accepts analog messages from the output port of the actuator and changes its state based on those messages
\end{enumerate}

@@

\Paragraph{Input and Output Ports}
Since the $\abftheory$-theory is a theory of agents, we consider ports as agents that
allows the exchange of information between a channel and another
agent.  However, a port must be considered as a special type of agent to avoid
an \emph{infinite regress} in which a port needs a port to transfer information
between the outside of the port to the inside of itself.

\begin{definition}{\bf Input or Output Port --}\label{def:port} 
	Ports are agents with the following predefined behavior: an input-port changes
	the type of information from assertion from a sender $s$ to an agent $a$,
	$\rassert{s}{a}$ to belief of the agent $a$ ($\beliefRegion_a$) while an
	output-port from belief to assertion; forwarding information from the outside
	of an agent's boundary to the inside (input-port) or vice versa (output-port). 
\end{definition}
The \emph{quality of a port} is determined by the $\rcc$ relation between the 
assertions received or sent and the belief, i.e. $\Rcc{\assertionRegion_s}{\beliefRegion_a}$ for the input-port
or $\Rcc{\assertionRegion_r}{\beliefRegion_a}$.

\begin{definition}{\bf Secure Input Port --}\label{def:secport}
	A secure input port allows information as incoming assertions to flow
	from a sender $s$ to the behavior of the recipient $a$ (agent).  If $p_I$ is a secure input port,
	$\world\models\varphi$, and $\varphi\in\assertionRegion_s$ where $s$
	has the only input port $p_I$, $\world'\models\beliefRegion_a$ and $\varphi\in\beliefRegion_a$.
\end{definition}

@@

\Paragraph{Communication Channels}
We start by considering the difference between a (communication)
mono-directional channel (channel from now on) and an agent, as we did for the
ports, since the $\abftheory$-theory is a theory of agents.  In fact, if a channel
were considered an agent (channel-agent) then the question would be how an
agent would transfer its Assertions to the channel-agent. If the channel
between the agent and the channel-agent is again an agent, we would generate an
\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
finite depth (of detail) for a channel, where there exists a bottom-channel
which is not an agent. For now, we do not constrain a channel-agent in any way
so there is no difference between a channel agent and agent. Therefore, for the
sake of simplicity, we consider, in this text, channels to be bottom-channels,
defined as agents with the pre-defined behavior (i.e. defined in a dogmatic
way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
prevents us from introducing additional constraints to the channel as storing
Assertions that are transferred over the channel, or filter out some
input-assertions.}.

@@

\begin{enumerate}[start=5, label={W\arabic*)}]
	\item \emph{secure input port} and \emph{output drop port} ($s\sdcom r$), 
	\item \emph{secure input port} and \emph{output insertion port} ($s\sicom r$), 
	\item \emph{secure input port} and \emph{output injection port} ($s\sjcom r$), 
	\item \emph{secure input port} and \emph{output selective drop port} ($s\ssdcom r$), 
	\item \emph{secure input port} and \emph{output selective insertion port} ($s\ssicom r$), 
	\item \emph{secure input port} and \emph{output selective injection port} ($s\ssjcom r$),

	\item \emph{input drop port} and \emph{output drop port} ($s\ddcom r$)
	\item \emph{input drop port} and \emph{output insertion port} ($s\dicom r$)
	\item \emph{input drop port} and \emph{output secure port} ($s\dscom r$)

	\item \emph{input insert port} and \emph{output drop port} ($s\idcom r$)
	\item \emph{input insert port} and \emph{output insert port} ($s\iicom r$)
	\item \emph{input insert port} and \emph{output injection port} ($s\ijcom r$)
	\item \emph{input insert port} and \emph{output selective drop port} ($s\isdcom r$)
	\item \emph{input insert port} and \emph{output selective insertion port} ($s\isicom r$)
	\item \emph{input insert port} and \emph{output selective injection port} ($s\isjcom r$)
	\item \emph{input insert port} and \emph{output secure port} ($s\iscom r$)

	\item \emph{input injection port} and \emph{output drop port} ($s\jdcom r$)
	\item \emph{input injection port} and \emph{output insertion port} ($s\jicom r$)
	\item \emph{input injection port} and \emph{output injection port} ($s\jjcom r$)
	\item \emph{input injection port} and \emph{output selective drop port} ($s\jsdcom r$)
	\item \emph{input injection port} and \emph{output selective insertion port} ($s\jsicom r$)
	\item \emph{input injection port} and \emph{output selective injection port} ($s\jsjcom r$)
	\item \emph{input injection port} and \emph{output secure port} ($s\jscom r$)

	\item \emph{input selective drop port} and \emph{output drop port} ($s\sddcom r$)
	\item \emph{input selective drop port} and \emph{output insertion port} ($s\sdicom r$)
	\item \emph{input selective drop port} and \emph{output injection port} ($s\sdjcom r$)
	\item \emph{input selective drop port} and \emph{output selective drop port} ($s\sdsdcom r$)
	\item \emph{input selective drop port} and \emph{output selective insertion port} ($s\sdsicom r$)
	\item \emph{input selective drop port} and \emph{output selective injection port} ($s\sdsjcom r$)
	\item \emph{input selective drop port} and \emph{output secure port} ($s\sdscom r$)

	\item \emph{input selective insertion port} and \emph{output drop port} ($s\sidcom r$)
	\item \emph{input selective insertion port} and \emph{output insertion port} ($s\siicom r$)
	\item \emph{input selective insertion port} and \emph{output injection port} ($s\sijcom r$)
	\item \emph{input selective insertion port} and \emph{output selective drop port} ($s\sisdcom r$)
	\item \emph{input selective insertion port} and \emph{output selective insertion port} ($s\sisicom r$)
	\item \emph{input selective insertion port} and \emph{output selective injection port} ($s\sisjcom r$)
	\item \emph{input selective insertion port} and \emph{output secure port} ($s\siscom r$)

	\item \emph{input selective injection port} and \emph{output drop port} ($s\sjdcom r$)
	\item \emph{input selective injection port} and \emph{output insertion port} ($s\sjicom r$)
	\item \emph{input selective injection port} and \emph{output injection port} ($s\sjjcom r$)
	\item \emph{input selective injection port} and \emph{output selective drop port} ($s\sjsdcom r$)
	\item \emph{input selective injection port} and \emph{output selective insertion port} ($s\sjsicom r$)
	\item \emph{input selective injection port} and \emph{output selective injection port} ($s\sjsjcom r$)
	\item \emph{input selective injection port} and \emph{output secure port}  ($s\sjcom r$)
\end{enumerate}
In Appendix~\ref{app:insecurecom}, we report the proof which is exhaustive on all
possible combinations.

\Paragraph{Functional Block}
A Functional Architecture, as already described beforehand, takes Information as input-Beliefs
and transforms the Information into output-Beliefs. Those transformations 
occurs within the Functional Architecture, where Functional Blocks transforms Beliefs into 
other Beliefs. Similarly to channels, we could consider a Functional Block as a Functional Architecture
occurring in an infinite regress. Therefore, we consider Functional Blocks as 
correctly executing their abstract behavior, correctly transforming inputs into outputs (Beliefs).

@@

\Paragraph{Facts}
During the specification phase, for any agent, channel, port, functional block
and architecture, there exist a Fact predicating over them.
In other words, any requirement is defined as a Fact since they must be true in
any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
Figure~\ref{fig:soundness}, Facts are strategic rules that defines how
the system shall behave (by specification), while reality may be shown to
be insecure.

The quality of a Functional Block is determined by the relation between the
input-Beliefs formulas and the symmetric formulas in the Fact region
(equivalently for output-Beliefs). Therefore, the quality of the functional
architecture is not a purely architectural property but relies on the logic
implemented in the functional blocks (which we abstract away at this stage of
the engineering model).
\begin{definition}{\bf Secure Functional Block and Architecture --}\label{def:secfun}
	A functional block is secure if the input beliefs and output beliefs
	are in $EQ$ relation with the corresponding Facts mandated by requirements.
	A functional architecture composed exclusively by secure functional blocks
	is considered secure.
\end{definition}

Therefore, there exists the following weaknesses in the functional
architecture, where $\beliefRegion_I$ and $\beliefRegion_O$ are input and
output Beliefs respectively, and $\factRegion_I$ and $\factRegion_O$ are
the corresponding Facts (i.e. what really happens).
\begin{enumerate}
	\item $\dr{\beliefRegion_I}{\factRegion_I}$ and
		$\dr{\beliefRegion_O}{\factRegion_O}$, the input beliefs are
		misinterpreted and this result in the complete alteration of the
		output beliefs
	\item \fix{mr}{TBD}
\end{enumerate}

\Paragraph{Security Requirements} 
While standards (such
as IEC62443-1-3\fixnote{mr}{check the correct version}) defines requirements as
``information in transit/at-rest should be encrypted'' we believe that, at
specification level, security requirements can be formulated as subcategories of
the well-known CIA triad (where CIA stands for Confidentiality, Integrity,
Availability).
Infosec, or Information Security, defines the security risks related to
information with the de-facto standard CIA triad, which is often criticized
\autocite{CIAcriticismCPS} for being too general or non-adequate (e.g. by
adding authenticity which is often used as a building block for confidentiality
and integrity) to be effectively applied to the engineering of systems. Due to
this, many researchers and organization tried to improve the CIA triad.  The
evolutions/extensions of the CIA triad has been documented, e.g., in
\autocite{Samonas2014cia} and is summarized in Table~\ref{tab:ciaevolution}.

>>>>

Several standards mandates a secure-by-design approach in which 
cybersecurity shall be considered at the very early stages of the 
design process. For example,
\begin{enumerate}[noitemsep]
	\item DO-326A -- ``Airworthiness Security Process Specification''
		requires a cybersecurity risk assessment of the design and is
		the ``are the only Acceptable Means of Compliance (AMC) by FAA
		\& EASA for aviation cyber-security airworthiness certification,
		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
		Control System (ICS) Security''
	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
		for Cyber-Physical Vehicle Systems'' defines ``set of
		high-level guiding principles for Cybersecurity as it relates
		to cyber-physical vehicle systems'' and states that
		``incorporate Cybersecurity into cyber-physical vehicle systems
		from concept phase through production, operation, service, and
		decommissioning''
\end{enumerate}
Companies who develop CPS (e.g. aerospace systems or elevator systems), must adhere
to strict regulations and quality assurances processes. Introducing tools and procedures 
in their processes requires a detailed and justified overview on how those
tools and procedures can be used. In Figure~\ref{fig:spdl}, we provide an overview
of a process that consider cybersecurity in the specification, design, and implementation
stages. The overall process start with the specification of the architectural (both physical
and functional) requirements from which the Weaknesses are automatically identified.
Based on the Assets, architectures, and the Weaknesses the cybersecurity risk is
calculated based on the $\abftheory$-theory. An iterative process between 
the user (e.g. the engineer) and the \secramod{} module of the \abftool{} allows the user to move to the design phase
when the risk level is considered acceptable. The design phase starts by automatically mapping
the Weaknesses into Vulnerabilities which are, in turn, fed into the \designverifmod{}
module along with the final specification. The Module returns a list of Vulnerabilities and
potential Mitigations. The user then procees in implementing the CPS and the SW/HW choices
can be tested by the \atgmod. \fix{mr}{While we have described the overall view, from
specificaiton to implementation, in this text we only consider the process from specificaiton to design.}

@@

In order to explain in details the SPDL, we use the following running example.
\begin{example}
	As depicted in Figure~\ref{fig:usecase}, the use case of the running example
	(a small part of a subprocess of the SwAT testbed \autocite{Mathur2016swat}), shows
	\begin{itemize}[noitemsep]
		\item a tank that is filled with raw water coming from the inlet pipe 
		\item a motorized valve (a valve with an actuator) such that:
			\begin{itemize}
				\item if opened, allows the raw water to flow into the tank
				\item if closed, stops the raw water to flow into the tank
			\end{itemize}
		\item a sensor that reads the meters of raw water in the tank
		\item a PLC controller, connected to the sensor and the
			actuator, that receives the readings from the sensor
			and regulates the quantity of water in the tank by
			communicating a change in the state of the actuator
			based on the following logic
			\begin{itemize}
				\item if the readings from the sensor reports a
					water level in the tank above (or
					equal) to 10m, the actuator shall close
					the valve
				\item if the readings are below 10m the actuator
					shall open the valve.
			\end{itemize}
	\end{itemize}

@@

\subsection{Specification -- Elicitation of Security Requirements}\label{sec:properties}
\begin{figure}[t]
	\centering
	\includegraphics[width=\textwidth]{spec.pdf}
	\caption{Specification -- Running Example}
	\label{fig:spec}
\end{figure}
The first step of the SPDL is the definition of the requirements and then, for
the purpose of this paper, of (i) security requirements, (ii) architectural
(physical and functional) requirements.  Our focus in mainly on security
requirements but they are necessarily intertwined with the architectural
requirements, as we will show afterwards in this section.  A complete study on
how to properly define the architectural requirements is outside the scope of
this paper so we propose a method that is convenient for the purpose of our
assessment.

\Paragraph{System Requirements}
In our running example, there is a system (a CPS) composed by $5$ main
subsystem which we consider atomic (i.e. not themself composed by subsystems)
and then agents: the tank, the sensor, the controller, the actuator, and the
motorized valve (valve from now on).  The whole specification (which we now
describe) of the physical and functional architecture is depicted in
Figure~\ref{fig:spec}.  In the use case, the communication flow is as follows:
\begin{enumerate}
	\item the sensor reads the level of water in the tank ($\text{tank}\rightarrow \text{sensor}$) \footnote{this communication depends on the technology in use for the readings. We assume a unidirectional communication without loss of generality}
	\item the sensor communicates the readings to the controller ($\text{sensor}\rightarrow \text{controller}$)
	\item the controller calculates the state of the valve and communicates it to the actuator ($\text{controller}\rightarrow \text{actuator}$)
	\item the actuator translates (digital to analog) the state received and communicates\footnote{we do not distinguish, for the sake of simplicity between different types of channels and assume that communicating to the valve produces the expected change of physical state} it to the valve
\end{enumerate}
Each agent has an input or output port from which it communicates with other
agents.  It is also composed by the functional blocks that defines its
functional architecture and physical boundaries that defines its physical
architecture, as follows.
\begin{enumerate}[start=1, label={req\arabic*)}]
	\item the tank has no functional architecture, it is a purely physical component which has the only purpose of containing water
	\item the communication is unidirectional from the tank to the input port of the sensor
	\item the sensor has an input-port that receives only incoming communication from the tank
	\item the sensor has two functional blocks
		\begin{enumerate}[label={req4.\arabic*)}]
			\item read\_water\_level, accepts inputs only from the input-port and outputs a new reading every $t$ seconds
			\item comm\_water\_level, accepts inputs only from the read\_water\_level and communicates the readings to the output-port
		\end{enumerate}
	\item the sensor has an output-port that only outputs the messages generated by the comm\_water\_level to the input-port of the controller
	\item the controller has in input-port that only receives incoming communications from the sensor and sends the messages to the functional architecture
	\item the functional architecture of the controller is composed by three functional blocks
		\begin{enumerate}[label={req7.\arabic*)}]
			\item get\_water\_level, accepts inputs only from the input-port and outputs the water level to the main control function
			\item control, calculate the next state of the valve based on the inputs receive from the get\_water\_level
			\item send\_state, receives the output of the control function and communicates it to the output-port
		\end{enumerate}
	\item the controller has an output-port that only outputs the messages generated by the send\_state to the input-port of the actuator 
	\item the actuator has an input-port that receives only incoming communication from the controller 
	\item the functional architecture of the actuator is composed by two functional blocks
		\begin{enumerate}[label={req10.\arabic*)}]
			\item read\_change\_state, accepts inputs only from the input-port and outputs the next state of the valve to the close\_open\_valve function
			\item close\_open\_valve, converts the digital next state received from the read\_change\_state an analog signal and communicates it the output-port
		\end{enumerate}
	\item the valve accepts analog messages from the output port of the actuator and changes its state based on those messages
\end{enumerate}

@@

\subsubsection{Input and Output Ports}
Since the $\abftheory$-theory is a theory of agents, we consider ports as agents that
allows the exchange of information between a channel and another
agent.  However, a port must be considered as a special type of agent to avoid
an \emph{infinite regress} in which a port needs a port to transfer information
between the outside of the port to the inside of itself.

\begin{definition}{\bf Input or Output Port --}\label{def:port} 
	Ports are agents with the following predefined behavior: an input-port changes
	the type of information from assertion from a sender $s$ to an agent $a$,
	$\rassert{s}{a}$ to belief of the agent $a$ ($\beliefRegion_a$) while an
	output-port from belief to assertion; forwarding information from the outside
	of an agent's boundary to the inside (input-port) or vice versa (output-port). 
\end{definition}
The \emph{quality of a port} is determined by the $\rcc$ relation between the 
assertions received or sent and the belief, i.e. $\Rcc{\assertionRegion_s}{\beliefRegion_a}$ for the input-port
or $\Rcc{\assertionRegion_r}{\beliefRegion_a}$. We only define secure input port but the
definition of a secure output port is symmetrical.

\begin{definition}{\bf Secure Input Port --}\label{def:secport}
	A secure input port ($s\secureport a$) allows information as incoming assertions to flow
	from a sender $s$ to the behavior of the recipient $a$ (agent).  If $p_I$ is a secure input port,
	$\world\models\varphi$, $\varphi\in\rassert{s}{a}$, and $a$
	has the only input port $p_I$, $\world'\models\beliefRegion_a$ and $\varphi\in\beliefRegion_a$ for any
	$\world\modalrelation\world'$.
\end{definition}

@@

\subsubsection{Communication Channels}
We start by considering the difference between a (communication)
mono-directional channel (channel from now on) and an agent, as we did for the
ports, since the $\abftheory$-theory is a theory of agents.  In fact, if a channel
were considered an agent (channel-agent) then the question would be how an
agent would transfer its Assertions to the channel-agent. If the channel
between the agent and the channel-agent is again an agent, we would generate an
\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
finite depth (of detail) for a channel, where there exists a bottom-channel
which is not an agent. For now, we do not constrain a channel-agent in any way
so there is no difference between a channel agent and agent. Therefore, for the
sake of simplicity, we consider, in this text, channels to be bottom-channels,
defined as agents with the pre-defined behavior (i.e. defined in a dogmatic
way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
prevents us from introducing additional constraints to the channel as storing
Assertions that are transferred over the channel, or filter out some
input-assertions.}.

@@

\begin{enumerate}[start=5, label={W\arabic*)}]
	\item \emph{secure output port} and \emph{input drop port} ($s\sdcom r$), 
	\item \emph{secure output port} and \emph{input insertion port} ($s\sicom r$), 
	\item \emph{secure output port} and \emph{input injection port} ($s\sjcom r$), 
	\item \emph{secure output port} and \emph{input selective drop port} ($s\ssdcom r$), 
	\item \emph{secure output port} and \emph{input selective insertion port} ($s\ssicom r$), 
	\item \emph{secure output port} and \emph{input selective injection port} ($s\ssjcom r$),

	\item \emph{output drop port} and \emph{input drop port} ($s\ddcom r$)
	\item \emph{output drop port} and \emph{input insertion port} ($s\dicom r$)
	\item \emph{output drop port} and \emph{input secure port} ($s\dscom r$)

	\item \emph{output insert port} and \emph{input drop port} ($s\idcom r$)
	\item \emph{output insert port} and \emph{input insert port} ($s\iicom r$)
	\item \emph{output insert port} and \emph{input injection port} ($s\ijcom r$)
	\item \emph{output insert port} and \emph{input selective drop port} ($s\isdcom r$)
	\item \emph{output insert port} and \emph{input selective insertion port} ($s\isicom r$)
	\item \emph{output insert port} and \emph{input selective injection port} ($s\isjcom r$)
	\item \emph{output insert port} and \emph{input secure port} ($s\iscom r$)

	\item \emph{output injection port} and \emph{input drop port} ($s\jdcom r$)
	\item \emph{output injection port} and \emph{input insertion port} ($s\jicom r$)
	\item \emph{output injection port} and \emph{input injection port} ($s\jjcom r$)
	\item \emph{output injection port} and \emph{input selective drop port} ($s\jsdcom r$)
	\item \emph{output injection port} and \emph{input selective insertion port} ($s\jsicom r$)
	\item \emph{output injection port} and \emph{input selective injection port} ($s\jsjcom r$)
	\item \emph{output injection port} and \emph{input secure port} ($s\jscom r$)

	\item \emph{output selective drop port} and \emph{input drop port} ($s\sddcom r$)
	\item \emph{output selective drop port} and \emph{input insertion port} ($s\sdicom r$)
	\item \emph{output selective drop port} and \emph{input injection port} ($s\sdjcom r$)
	\item \emph{output selective drop port} and \emph{input selective drop port} ($s\sdsdcom r$)
	\item \emph{output selective drop port} and \emph{input selective insertion port} ($s\sdsicom r$)
	\item \emph{output selective drop port} and \emph{input selective injection port} ($s\sdsjcom r$)
	\item \emph{output selective drop port} and \emph{input secure port} ($s\sdscom r$)

	\item \emph{output selective insertion port} and \emph{input drop port} ($s\sidcom r$)
	\item \emph{output selective insertion port} and \emph{input insertion port} ($s\siicom r$)
	\item \emph{output selective insertion port} and \emph{input injection port} ($s\sijcom r$)
	\item \emph{output selective insertion port} and \emph{input selective drop port} ($s\sisdcom r$)
	\item \emph{output selective insertion port} and \emph{input selective insertion port} ($s\sisicom r$)
	\item \emph{output selective insertion port} and \emph{input selective injection port} ($s\sisjcom r$)
	\item \emph{output selective insertion port} and \emph{input secure port} ($s\siscom r$)

	\item \emph{output selective injection port} and \emph{input drop port} ($s\sjdcom r$)
	\item \emph{output selective injection port} and \emph{input insertion port} ($s\sjicom r$)
	\item \emph{output selective injection port} and \emph{input injection port} ($s\sjjcom r$)
	\item \emph{output selective injection port} and \emph{input selective drop port} ($s\sjsdcom r$)
	\item \emph{output selective injection port} and \emph{input selective insertion port} ($s\sjsicom r$)
	\item \emph{output selective injection port} and \emph{input selective injection port} ($s\sjsjcom r$)
	\item \emph{output selective injection port} and \emph{input secure port}  ($s\sjcom r$)
\end{enumerate}
In Appendix~\ref{app:insecurecom}, we report the proof which is exhaustive on all
possible combinations.

\subsubsection{Functional Block}
A Functional Architecture, as already described beforehand, takes Information as input-Beliefs
and transforms the Information into output-Beliefs. Those transformations 
occurs within the Functional Architecture, where Functional Blocks transforms Beliefs into 
other Beliefs. Similarly to channels, we could consider a Functional Block as a Functional Architecture
occurring in an infinite regress. Therefore, we consider Functional Blocks as 
correctly executing their abstract behavior, correctly transforming inputs into outputs (Beliefs).

@@

\Paragraph{Requirements as Facts}
During the specification phase, for any agent, channel, port, functional block
and architecture, there exist a Fact predicating over them.
In other words, any requirement is defined as a Fact since they must be true in
any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
Figure~\ref{fig:soundness}, Facts are strategic rules that defines how
the system shall behave (by specification), while reality may be shown to
be insecure.

The quality of a Functional Block is determined by the relation between the
input-Beliefs formulas and the symmetric formulas in the Fact region
(equivalently for output-Beliefs).

\begin{definition}{\bf Secure Functional Block and Architecture --}\label{def:secfun}
	A functional block is secure if the input beliefs and output beliefs
	are in $EQ$ relation with the corresponding Facts mandated by requirements.
	A functional architecture composed exclusively by secure functional blocks
	is considered secure.
\end{definition}

Therefore, there exists the following weaknesses in the functional
architecture, where $\beliefRegion_I$ and $\beliefRegion_O$ are input and
output Beliefs respectively, and $\factRegion_I$ and $\factRegion_O$ are
the corresponding Facts (i.e. what really happens). 
\begin{itemize}
	\item $\eq{\beliefRegion_I}{\factRegion_I}$ and then only specified
		inputs can be fed to the functional block but:
	\begin{enumerate}[start=49, label={W\arabic*)}]
		\item $\dr{\beliefRegion_O}{\factRegion_O}$, not all the specified
			outputs are produced
		\item $\po{\beliefRegion_O}{\factRegion_O}$ , only some but not all of the specified
			outputs are produced along with some non-specified output are produced
		\item $\pp{\beliefRegion_O}{\factRegion_O}$ , not all the specified outputs are produced 
		\item $\ppi{\beliefRegion_O}{\factRegion_O}$ , not all the output have been specified
	\end{enumerate}
	\item $\po{\beliefRegion_I}{\factRegion_I}$, not all the specified inputs can be fed into the functional block
	\begin{enumerate}[start=53, label={W\arabic*)}]
		\item $\dr{\beliefRegion_O}{\factRegion_O}$
		\item $\po{\beliefRegion_O}{\factRegion_O}$
		\item $\pp{\beliefRegion_O}{\factRegion_O}$
		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
		\item $\eq{\beliefRegion_O}{\factRegion_O}$ this case implies
			that somehow there is no input that generates an
			non-specified output but this is not because the inputs
			are secure 
	\end{enumerate}
	\item $\pp{\beliefRegion_I}{\factRegion_I}$, not all the specified
		inputs can be fed into the block
	\begin{enumerate}[start=58, label={W\arabic*)}]
		\item $\dr{\beliefRegion_O}{\factRegion_O}$
		\item $\po{\beliefRegion_O}{\factRegion_O}$
		\item $\pp{\beliefRegion_O}{\factRegion_O}$
		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
		\item $\eq{\beliefRegion_O}{\factRegion_O}$
	\end{enumerate}
	\item $\ppi{\beliefRegion_I}{\factRegion_I}$, not all the inputs that
		can be fed into the functional block follows the specification
	\begin{enumerate}[start=63, label={W\arabic*)}]
		\item $\dr{\beliefRegion_O}{\factRegion_O}$
		\item $\po{\beliefRegion_O}{\factRegion_O}$
		\item $\pp{\beliefRegion_O}{\factRegion_O}$
		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
		\item $\eq{\beliefRegion_O}{\factRegion_O}$
	\end{enumerate}
	\item $\dr{\beliefRegion_I}{\factRegion_I}$, no specified inputs can be
		fed into the functional block
	\begin{enumerate}[start=68, label={W\arabic*)}]
		\item $\dr{\beliefRegion_O}{\factRegion_O}$
		\item $\po{\beliefRegion_O}{\factRegion_O}$
		\item $\pp{\beliefRegion_O}{\factRegion_O}$
		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
		\item $\eq{\beliefRegion_O}{\factRegion_O}$
	\end{enumerate}
\end{itemize}

\subsubsection{Security Requirements} 
While standards (such
as IEC62443-1-3\fixnote{mr}{check the correct version}) defines requirements as
``information in transit/at-rest should be encrypted'' we believe that, at
specification level, security requirements can be formulated as subcategories of
the well-known CIA triad (where CIA stands for Confidentiality, Integrity,
Availability).
Infosec, or Information Security, defines the security risks related to
information with the de-facto standard CIA triad, which is often criticized
\autocite{CIAcriticismCPS} for being too general or non-adequate (e.g. by
adding authenticity which is often used as a building block for confidentiality
and integrity) to be effectively applied to the engineering of systems. Due to
this, many researchers and organization tried to improve the CIA triad.  The
evolutions/extensions of the CIA triad has been documented, e.g., in
\autocite{Samonas2014cia} and is summarized in Table~\ref{tab:ciaevolution}.
